---
title: "PA1_template"
author: "Chris"
date: "5/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,output = FALSE)
library(ggplot2)
library(dplyr)
library(scales)
library(lubridate)
setwd('C:/Users/chris/OneDrive/Desktop/repdata_data_activity/RepData_PeerAssessment1')
```

### Loading and preprocessing the data

Here I have:

    - Read in the data
    
    - Counted the number of missing values with the complete.cases function
    
    - Changed the date variable to a date format

```{r initial read}
activity<-read.csv('activity/activity.csv')
activity$date<-ymd(activity$date)
```
### First Histogram: Total Number of Steps per Day

```{r histogram}
activity %>%
    group_by(date) %>%
    summarize(sum_steps=sum(steps)) %>%
    ggplot(aes(x=date,y=sum_steps)) + 
    geom_bar(stat='identity') + 
    scale_x_date(breaks=date_breaks('weeks'))
```
### Daily Means and Medians with Original Data

```{r mean and median}
activity %>%
    group_by(date) %>%
    summarize(mean_steps=mean(steps,na.rm=T),median_steps=median(steps,na.rm=T))
```
### Time Series Plot of Average Steps over Intervals

```{r time series plot}
activity %>%
    group_by(interval) %>%
    summarize(avg_steps=mean(steps,na.rm=T)) %>%
    ggplot(aes(x=interval)) +
    geom_line(aes(y=avg_steps))
```
```{r max interval}
max.values<-activity %>%
    group_by(interval) %>%
    summarize(max_steps=max(steps,na.rm=T))
max.set<-max.values[which(max.values$max_steps==max(max.values$max_steps)),]
print(paste0('The interval with the highest average of daily steps is ',max.set$interval,' with an average of ',
             max.set$max_steps,' steps'))
```

### Addressing Missing Values

```{r replacing missing values}
x<-complete.cases(activity)
NA.count<-length(activity[,1])-sum(x)
print(paste0('Number of NAs in dataset = ',NA.count))
```

Here I have:

    - Created a new dataset called NAs.replaced
    
    - In the new dataset, I replaced the missing values in the step variable 
        with the mean of that interval's steps across all day
    
```{r}
replace<-activity %>%
    group_by(interval) %>%
    mutate(mean_steps=mean(steps,na.rm=T))
NAs.replaced<-activity
for(i in 1:length(NAs.replaced$steps)){
    if(is.na(NAs.replaced$steps[i])){
        NAs.replaced$steps[i]<-replace$mean_steps[i]
    }
}
```

First 3 lines of activity dataset:

```{r}
activity[1:3,]
```

First 3 lines of activity dataset:

```{r}
NAs.replaced[1:3,]
```
### Second Histogram: Total Number of Steps per Day with NAs Removed

```{r}
NAs.replaced %>%
    group_by(date) %>%
    ggplot(aes(date,steps))+
    stat_summary(fun=sum,geom='bar')+
    ggtitle('Daily Sums of Steps')
```

### Daily Means and Medians with NAs Removed

The main difference is that there are no NaN values.

```{r}
NAs.replaced %>%
    group_by(date) %>%
    summarize(mean_steps=mean(steps,na.rm=T),median_steps=median(steps,na.rm=T))
```

Adding 2 new variables to the dataset with the NAs removed

    DayofWeek (Day of the Week) and TimeofWeek (Weekday or Weekend)

```{r}
NAs.replaced<-NAs.replaced %>%
    mutate(DayofWeek=wday(date,label=T)) %>%
    mutate(TimeofWeek=ifelse(DayofWeek%in%c('Sat','Sun'),'Weekend','Weekday'))
NAs.replaced$TimeofWeek<-as.factor(NAs.replaced$TimeofWeek)
```
### Time Series Plot Split by Time of Week

Splitting the plot by the factor of TimeofWeek

```{r split time series plot}
NAs.replaced %>%
    group_by(TimeofWeek,interval) %>%
    summarize(avg_steps=mean(steps,na.rm=T)) %>%
    ggplot(aes(x=interval)) +
    geom_line(aes(y=avg_steps)) + 
    facet_wrap(~TimeofWeek)
```


